Random Walks across Social Networks: The Markovian Movement of a Stochastic Soccerball

### 2020-07-20 rmd/draft.Rmd
almost done writing

I used Python to convert this into a *tidy* dataframe [@HWickham2014], that could then be analyzed and visualized using R. For the sake of reproducible research, all code is attached as appendices, as well as available on Github.

@@

In Figure 8 we show the heatmap of the Transition Matrix derived from the Adjacency Matrix shown in Figure 6. There we showed only the adjacency counts for the starting eleven, but now we include the subsitutes, as well as treating the 'Opposition' collectively as a single node. Most significantly, we now include a 'Goal' as a node. In every other case where the ball goes out of play, we ignore the details of whether it was for a corner, throw-in, offside, or foul, and simply consider the ball to have moved from whichever player had the ball to the opposition. (Consequently, the players which the Adjacency Matrix indicates get the ball most often from the opposition are generally full-backs who take throw-ins, rather than players whose challenges are actually winning the ball, as one might initially suppose). But a goal is of unique significance, and for a social network analysis of football to be anything more than a diverting curiosity, we need to be able to say something meaningful about what a team's possession graph tells us about the likelihood of goals.

To do this, as we have said, we treat a goal as a node in its own right; in particular, in the terms of Markov theory, we treat it as an *absorbing state* which loops back on itself with certain probability. Clearly a football match continues after a goal is scored, but from the perspective of our model, the match is considered as a series of stochastic processes which end and immediately start anew if and when a goal is scored. This allows us to calculate the expected *hitting time* of a goal, given that any particular player has the ball.

@@

## Results

T

@@

@@

>>>>

I used Python to convert this into a *tidy* dataframe [@HWickham2014], which I then analyzed and visualized using R. For the sake of reproducible research [@RPeng2011], all code is attached as appendices, as well as available on Github.

@@

In Figure 8 we show the heatmap of the Transition Matrix derived from the Adjacency Matrix shown in Figure 6. There we showed only the adjacency counts for the starting eleven, but now we include the subsitutes, as well as treating the 'Opposition' collectively as a single node. Most significantly, we now include a 'Goal' as a node. In particular, in the terms of Markov theory, we treat it as an *absorbing state* which loops back on itself with certain probability. This allows us to calculate the expected *hitting time* of a goal, given that any particular player has the ball.

@@

### Results

The

@@

We have here focussed on the attacking contribution a player makes towards his team's goal-scoring capability, but we could equally consider the defensive aspect by letting the absorbing state be the possibility that the opposition score.

One apparent flaw of this analysis is that it pays no attention to the *spatial* dynamics of the game. But it would be fairly straightforward to develop the analysis articulated here to also take into account a player's location on the pitch when he plays the ball. Rather than simply representing the players, the nodes' state space could instead be the Cartesian product of players with some set of discrete areas of the pitch, say $\{\Delta, \Phi, \Xi \}$ could represent the Defensive Half, Attacking Half, and Attacking Penalty Area respectively. One could simply add the spatial indicator to the player name at the preprocessing stage, and the rest of the analysis would then proceed as described here.

But the major weakness is one inherent in the data itself -- although the event-log dataset offers a wealth of information when compared to the manual *notational analysis* first developed by Charles Reep in the 1950s [@RPollard2002], a convincing analysis of football will only really be possible with data that includes the positions of off-the-ball players, as well as on-the-ball events. Such data can be captured [eg. @SPettersenEtAl2014], but is not yet widely (at least openly) available.

## Acknowledgments

@@

## Appendix 1: Analysis in R

## Appendix 2: Preprocessing in Python
